---
title: "Homework 2"
author: Na Yun Cho
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

# Problem 1

## Clean the Mr. Trash Wheel dataset
```{r}
trashweel_df =
  read_xlsx("./data/trashwheel_data.xlsx", 
          sheet = "Mr. Trash Wheel",
          range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls))

```


# Problem 2

## Clean the NYC Transit dataset, convert variable types
```{r}
transit_df = read.csv("./data/transit_data.csv") %>%
  janitor::clean_names() %>%
  select(c(line:entry, vending, ada)) %>%
  mutate(entry = recode(entry, 'YES' = TRUE, 'NO' = FALSE)) 

```

Short paragraph about the dataset: 
This dataset shows information regarding each entrance and exit for each subway station in New York City.
Some of the imporant variables present in the dataset are 'Station Name', 'Line', 'Routes', 'Vending', 'ADA', 'Entry', and 'Entrance Type'.
The dataset has `r nrow(transit_df)` rows and `r ncol(transit_df)` columns. 
In order to clean the data, so far I have attempted to format the variable names more consistently and selected the columns that I need most. 
I also turned the 'entry' variable from character type to a logical variable. 
However, the data in the dataset are not tidy yet because the route number and route name are still kept as the same variable.

Answer 1 of the dataset: 
there are `r nrow(distinct(transit_df, line)) * nrow(distinct(transit_df, station_name))` distinct stations.

## Stations that are ADA compliant
```{r}
transit_df_ada = filter(transit_df, ada == TRUE)
```

Answer 2 of the dataset:
there are `r nrow(distinct(transit_df_ada, line)) * nrow(distinct(transit_df_ada, station_name))` stations that are ADA compliant.

## Station entrnaces/exits without vending
```{r}
transit_df_novend = filter(transit_df, vending != 'YES')
```

## Station entrances/exists without vending that allow entrance
```{r}
transit_df_novend_entry = filter(transit_df_novend, entry == TRUE)
```

Answer 3 of the dataset:
`r round(nrow(transit_df_novend_entry)/nrow(transit_df_novend) * 100, 2)`% of station entrances/exits without vending allow entrance.

## Reformat the data
```{r}
transit_df_reformatted = 
  transit_df %>%
  mutate(
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)) %>%
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route_name")
  
```

## Distinct stations that serve the A train
```{r}
transit_df_A = filter(transit_df_reformatted, route_name == 'A')
```

Answer 4 of the dataset: 
There are `r nrow(distinct(transit_df_A, line)) * nrow(distinct(transit_df_A, station_name))` distinct stations that serve the A train.

## ADA compliant stations that serve the A train
```{r}
transit_df_A_ada = filter(transit_df_A, ada == TRUE)
```

Answer 5 of the data set:
Of the stations that serve the A train, `r nrow(transit_df_A_ada)` are ADA compliant.




# Problem 3

## Clean the data in pols-month.csv 
```{r}
pols_month = read.csv("./data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c('year', 'month', 'day'), sep = '-') %>%
  mutate(month = as.numeric(month)) %>%
  mutate(month = month.abb[month]) %>%
  mutate(year = as.integer(year)) %>%
  mutate(president = case_when(
    prez_gop == 0 ~ 'dem',
    prez_gop == 1 ~ 'gop',
    prez_gop == 2 ~ 'dem')) %>%
  select(-c(prez_gop, prez_dem, day)) %>%
  relocate(year)

```

## Clean the data in snp.csv
```{r}
snp_df = read.csv("./data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c('month', 'day', 'year'), sep = '/') %>%
  mutate(month = as.numeric(month)) %>%
  mutate(month = month.abb[month]) %>%
  mutate(year = as.integer(year)) %>%
  relocate(year)
```

## Clean the data in unemployment.csv
```{r}
unemployment_df = read.csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment") %>%
  janitor::clean_names()
  
```

## Join the datasets
```{r}
snp_pols_df = left_join(pols_month, snp_df, by = c("year", "month")) 

final_df = left_join(snp_pols_df, unemployment_df, by = c("year", "month"))
                                                  
```


Short paragraph about these datasets:

First, the pol-month dataset contains information about the number of politicians who are democratic or republican at any given time. 
Second, the snp dataset shows information regarding the Standard & Poor's stock market index on any given date.
Lastly, the unemployment dataset contains information about the percentages of unemployment in any given month of the associated years. 
The final resulting dataset contains all the combined information from these three individual datasets merged together by the common variables of 'year' and 'month'. 
The dimensions of the resulting dataset are `r nrow(final_df)` rows and `r ncol(final_df)` columns. 
The range of the years covered in this final dataset is `r max(final_df$year)` - `r min(final_df$year)` years, which is 68 years.  
This dataset contains the following variables - `r ls(final_df)`. 
Among these variables, some of the key variables are 'year', 'month', 'president', 'unemployment', and 'close'.