---
title: "Homework 2"
author: Na Yun Cho
output: github_document
---

```{r}
library(tidyverse)
```

# Problem 2

# Clean the data, convert variable types
```{r}
transit_df = read.csv("./data/transit_data.csv") %>%
  janitor::clean_names() %>%
  select(c(line:entry, vending, ada)) %>%
  mutate(entry = recode(entry, 'YES' = TRUE, 'NO' = FALSE)) 

```

Short paragraph about the dataset: 



Answer 1 of the dataset: 
there are `r nrow(distinct(transit_df, line)) * nrow(distinct(transit_df, station_name))` distinct stations.

# Stations that are ADA compliant
```{r}
transit_df_ada = filter(transit_df, ada == TRUE)
transit_df_ada
```

Answer 2 of the dataset: 
there are `r nrow(distinct(transit_df_ada, line)) * nrow(distinct(transit_df_ada, station_name))` stations that are ADA compliant.

# Station entrnaces/exits without vending
```{r}
transit_df_novend = filter(transit_df, vending != 'YES')
```

# Station entrances/exists without vending that allow entrance
```{r}
transit_df_novend_entry = filter(transit_df_novend, entry == TRUE)
```

Answer 3 of the dataset: 
`r round(nrow(transit_df_novend_entry)/nrow(transit_df_novend) * 100, 2)`% of station entrances/exits without vending allow entrance.

# Reformat the data
```{r}
transit_df_reformatted = 
  transit_df %>%
  mutate(
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)) %>%
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route_name")
  
```

# Distinct stations that serve the A train
```{r}
transit_df_A = filter(transit_df_reformatted, route_name == 'A')
```

# ADA compliant stations that serve the A train
```{r}
transit_df_A_ada = filter(transit_df_A, ada == TRUE)
```

Answer 4 & 5 of the data set:
There are `r nrow(distinct(transit_df_A, line)) * nrow(distinct(transit_df_A, station_name))` distinct stations that serve the A train.
Of the stations that serve the A train, `r nrow(transit_df_A_ada)` are ADA compliant.


# Problem 3

# Clean the data in pols-month.csv 
```{r}
pols_month = read.csv("./data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c('year', 'month', 'day'), sep = '-') %>%
  mutate(month = as.numeric(month)) %>%
  mutate(month = month.abb[month]) %>%
  mutate(year = as.integer(year)) %>%
  mutate(president = case_when(
    prez_gop == 0 ~ 'dem',
    prez_gop == 1 ~ 'gop',
    prez_gop == 2 ~ 'dem')) %>%
  select(-c(prez_gop, prez_dem, day)) %>%
  relocate(year)

```

# Clean the data in snp.csv
```{r}
snp_df = read.csv("./data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c('month', 'day', 'year'), sep = '/') %>%
  mutate(month = as.numeric(month)) %>%
  mutate(month = month.abb[month]) %>%
  mutate(year = as.integer(year)) %>%
  relocate(year)
```

# Clean the data in unemployment.csv
```{r}
unemployment_df = read.csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment") %>%
  janitor::clean_names()
  
```

# Join the datasets
```{r}
snp_pols_df = left_join(pols_month, snp_df, by = c("year", "month")) 

final_df = left_join(snp_pols_df, unemployment_df, by = c("year", "month"))
                                                  
```

